---
title: "MA684 Midterm Project: Airbnb Data Analysis"
author: "Shuyi Jiang"
date: "12/03/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(sqldf)
library(tm)
library(SnowballC)
library(wordcloud)
library(RColorBrewer)
library(ggmap)
library(lme4)
library(Matrix)
library(knitr)
library(sjPlot)
```
# Project overview and goal
As a beginner who is new to airbnb, I really have no idea how to find an ideal place to stay on this platform. I looked up the beginner guides online and figured out that what the majority of potential houseguests care the most are price, location and host. How people choose airbnb should be an interesting angle for marketing. The project will focus on how to find a nice airbnb at a good price, in other words, to figure out the relationship bewteen price and factors that houseguests care the most. 

# Boston: data cleaning
I choosed Boston, the city I am the most familiar with, for this project. The data includes listing information in Boston from January to November in 2016. The sample size is 36359 (1 rows was removed because of missing data).
```{r}
boston1 <- read.csv("boston1.csv")
boston2 <- read.csv("boston2.csv")
boston3 <- read.csv("boston3.csv")
boston4 <- read.csv("boston4.csv")
boston5 <- read.csv("boston5.csv")
boston6 <- read.csv("boston6.csv")
boston7 <- read.csv("boston7.csv")
boston8 <- read.csv("boston8.csv")
boston9 <- read.csv("boston9.csv")
boston10 <- read.csv("boston10.csv")
boston11 <- read.csv("boston11.csv")

boston1$month <- 1
boston2$month <- 2
boston3$month <- 3
boston4$month <- 4
boston5$month <- 5
boston6$month <- 6
boston7$month <- 7
boston8$month <- 8
boston9$month <- 9
boston10$month <- 10
boston11$month <- 11

boston <- rbind(boston1,boston2,boston3,boston4,boston5,
                boston6,boston7,boston8,boston9,boston10,boston11)
boston <- boston[-c(4,14)]
boston <- filter(boston, !is.na(host_id))
```

# Word cloud
Looking at the word clouds of top 40 words about listing in name (searching key words), summary and description, I figured out that houseguests pay the most attention of price, location and room type.
```{r}
datatext<-read.csv("listings.csv",stringsAsFactors = FALSE)
jeopCorpus <- Corpus(VectorSource(datatext$name))
jeopCorpus <- tm_map(jeopCorpus, PlainTextDocument)
jeopCorpus <- tm_map(jeopCorpus, stripWhitespace)
jeopCorpus <- tm_map(jeopCorpus, tolower)
jeopCorpus <- tm_map(jeopCorpus, removeNumbers)
jeopCorpus <- tm_map(jeopCorpus, removePunctuation)
jeopCorpus <- tm_map(jeopCorpus, removeWords, stopwords('english'))
jeopCorpus <- tm_map(jeopCorpus, stemDocument)
jeopCorpus <- tm_map(jeopCorpus, removeWords, "bedroom")
jeopCorpus <- tm_map(jeopCorpus, removeWords, "room")
jeopCorpus <- tm_map(jeopCorpus, removeWords, "boston")
pal<-brewer.pal(10,"Dark2")
wordcloud(jeopCorpus, max.words = 40, random.order = FALSE,colors=pal)
jeopCorpus <- Corpus(VectorSource(datatext$summary))
jeopCorpus <- tm_map(jeopCorpus, PlainTextDocument)
jeopCorpus <- tm_map(jeopCorpus, stripWhitespace)
jeopCorpus <- tm_map(jeopCorpus, tolower)
jeopCorpus <- tm_map(jeopCorpus, removeNumbers)
jeopCorpus <- tm_map(jeopCorpus, removePunctuation)
jeopCorpus <- tm_map(jeopCorpus, removeWords, stopwords('english'))
jeopCorpus <- tm_map(jeopCorpus, stemDocument)
jeopCorpus <- tm_map(jeopCorpus, removeWords, "bedroom")
jeopCorpus <- tm_map(jeopCorpus, removeWords, "room")
jeopCorpus <- tm_map(jeopCorpus, removeWords, "boston")
pal<-brewer.pal(10,"Dark2")
wordcloud(jeopCorpus, max.words = 40, random.order = FALSE,colors=pal)
jeopCorpus <- Corpus(VectorSource(datatext$description))
jeopCorpus <- tm_map(jeopCorpus, PlainTextDocument)
jeopCorpus <- tm_map(jeopCorpus, stripWhitespace)
jeopCorpus <- tm_map(jeopCorpus, tolower)
jeopCorpus <- tm_map(jeopCorpus, removeNumbers)
jeopCorpus <- tm_map(jeopCorpus, removePunctuation)
jeopCorpus <- tm_map(jeopCorpus, removeWords, stopwords('english'))
jeopCorpus <- tm_map(jeopCorpus, stemDocument)
jeopCorpus <- tm_map(jeopCorpus, removeWords, "bedroom")
jeopCorpus <- tm_map(jeopCorpus, removeWords, "room")
jeopCorpus <- tm_map(jeopCorpus, removeWords, "boston")
pal<-brewer.pal(10,"Dark2")
wordcloud(jeopCorpus, max.words = 40, random.order = FALSE,colors=pal)
```

# Airbnb daily price in Boston
Based on what discussed about, I am interested in how location(neighborhood), room type and season affect price.
```{r}
summary(boston$price)
```
Only 769 out of 36359 samples have price over 500. Thus the following plots will only show samples with price under 500. The boxplots indicates that season (month) has very limited influence on price, while neighborhood and room type make relatively great influence on price.
```{r}
ggplot(boston, aes(x=month,y=price,group=month,colour=as.factor(month)))+
  geom_boxplot()+ylim(0,500)+labs(colour="month")
ggplot(boston, aes(x=neighborhood,y=price,group=neighborhood,colour=neighborhood))+
  geom_boxplot()+ylim(0,500)+theme(axis.text.x = element_text(angle = 45), text=element_text(size = 8))+
  labs(title="Airbnb price in Boston by neighborhood")
ggplot(boston, aes(x=room_type,y=price,group=room_type,colour=room_type))+
  geom_boxplot()+ylim(0,500)+labs(title="Airbnb price in Boston by room type")
```

# Room type of airbnb in Boston
There are 3 room types: entire home/apt, private room and shared room. Entire home/apt is the most popular room type (more than half of 36359 listings) and the average price of this type of room is more than twice of the other 2. Share room is only about 2% of all the listing airbnb and it has the lowest average price.
```{r}
room_type <- count(boston, room_type)
room_type
price_by_roomtype <- sqldf("SELECT room_type, avg(price) as avg_price 
                           FROM boston
                           GROUP BY room_type")
price_by_roomtype
ggplot(room_type,aes(x="",y=n,fill=room_type))+geom_bar(stat = "identity")+
  coord_polar("y")+labs(x="",y="",title="Pie Chart of Room Type of Airbnb in Boston")+
  geom_text(aes(label=n))
ggplot(price_by_roomtype, aes(x=room_type,y=avg_price, fill=room_type,colour=room_type))+
  geom_bar(stat = "identity")+geom_text(aes(label=round(avg_price, digits = 2)),vjust=-0.5)+
  theme(text=element_text(size = 10))+labs(title="Average airbnb price in Boston by room type")
```

# Neighborhood of airbnb in Boston
There are 25 different neighborhoods with airbnb listings in boston. The triangle area on the south riverbank of Charles River, including Allston, Fenway, Back Bay and South Boston is the most popular place for airbnb. This area does not have the lowest price but has relatively more convenient transportation.
```{r}
neighborhood <- count(boston, neighborhood)
neighborhood
price_by_neighborhood <- sqldf("SELECT neighborhood, avg(price) as avg_price 
                           FROM boston
                           GROUP BY neighborhood")
price_by_neighborhood
ggplot(neighborhood,aes(x="",y=n,fill=neighborhood))+geom_bar(stat = "identity")+
  coord_polar("y")+labs(x="",y="",title="Pie Chart of Neighborhood of Airbnb in Boston")+
  theme(text=element_text(size = 10))
ggplot(price_by_neighborhood, aes(x=neighborhood,y=avg_price, fill=neighborhood,
                                  colour=neighborhood))+
  geom_bar(stat = "identity")+theme(axis.text.x = element_text(angle = 90),
                                    text=element_text(size = 8))+
  geom_text(aes(label=round(avg_price, digits = 2)),vjust=-0.5,size=2,colour="black")+
  labs(title="Average airbnb price in Boston by neighborhood")
```

```{r}
mapboston <- (data.frame(
  x = boston$latitude,
  y = boston$longitude,
  month = boston$month
))
qmplot(y, x, data = mapboston, colour = as.factor(month), size = I(0.1), darken = .1)+
  labs(colour="month")
```

# Overall satisfication of airbnb in Boston
The listed airbnb in this data set is rated in a scale of 1 to 5 with 0.5 interval. The majority (9119 + 11224 = 20343) of 36359 samples is rated as 4.5 or 5 and 13203 samples, which is about 36%, are missing rating data. Thus, the overall satisfaction will be disregarded in the analysis because it is obviously not a representitive indicator.
```{r}
satisfaction <- count(boston, overall_satisfaction)
satisfaction
ggplot(satisfaction, aes(x=factor(overall_satisfaction),y=n, 
                         fill=overall_satisfaction,colour=overall_satisfaction))+
  geom_bar(stat = "identity")+geom_text(aes(label=n),vjust=-0.5)
```

# Model Testing
Multilevel linear model was choosen after several trials of different models. While fitting the multilevel, samples have NA in reviews, accommodates and bedrooms are removed.
```{r}
boston.model <- filter(boston, !is.na(accommodates) & !is.na(bedrooms) & !is.na(room_type))
```

Model fit1, fit2, fit3 and fit4 are varing-intercept models with predictors. They are using different combination of 3 variables (reviews, accommodates and bedrooms) to see whether it is better to include any of the 3 variables in the model. It turns out that fit1 with all 3 variables is the best among these 4 models.
```{r}
fit1 <- lmer(price ~ reviews + accommodates + bedrooms + 
               (1|neighborhood) + (1|room_type), REML = FALSE, data=boston.model)
summary(fit1)
```

```{r}
fit2 <- lmer(price ~ reviews + bedrooms + 
               (1|neighborhood) + (1|room_type), REML = FALSE, data=boston.model)
summary(fit2)
```

```{r}
fit3 <- lmer(price ~ reviews + accommodates + 
               (1|neighborhood) + (1|room_type), REML = FALSE, data=boston.model)
summary(fit3)
```

```{r}
fit4 <- lmer(price ~ accommodates + bedrooms + 
               (1|neighborhood) + (1|room_type), REML = FALSE, data=boston.model)
summary(fit4)
```

```{r}
anova(fit1,fit2,fit3,fit4)
```

# Model
The final model choosen is a multilevel model varying both intercepts and slopes.
$$y_{i}\sim\ N(\alpha_{j[i]}+\beta_{j[i]}x_{i},\ \sigma_{y}^2),\ for\ i=1,\dots,n$$
```{r}
fit5 <- lmer(price ~ reviews + accommodates + bedrooms +
               (1+reviews + accommodates + bedrooms|neighborhood) +
               (1+reviews + accommodates + bedrooms|room_type),
             REML = FALSE, data=boston.model)
summary(fit5)
```

As showen below, the coefficients of fitted model are quite significant, especially the number of rooms. Reviews (the number of total reviews of a specific room, which is used to predict the popularity of the this room) has much less influence than the other 2 variables. In general, with everything else remains the same, 1 increase in accommodates will lead to 4.2 higher daily price. One more room in airbnb will increase the daily price by 25.6 on average.
```{r}
kable(as.data.frame(fixef(fit5)))
```

```{r}
sjp.lmer(fit5, type = "re.qq")
```

```{r}
sjp.lmer(fit5, y.offset = .4)
```
The model has 2 multilevel predictors: neighborhood and room type. As shown in the graphs above, the influence of variables varies by different levels of predictors. For example, bedrooms increases the price more with the same change in neighborhood with higher price, such as Back bay.

```{r}
plot(fit5, ylim = c(-300,300), xlim = c(0,500))
```

# Discussion and concerns
The original data has lots of limitation, such as missing guest_id, which is necessary to do analysis of houseguests' review of rooms they have stayed in. Although the final model is better than the other testing models, it has huge AIC. Digging deeper with help of data other than the original data is definitely what I need to do after this first step of research. More variables that will make the model better has to be determined.
